%%%%Randomize
startdef insvar = int(30*random.choice([.7,.8,.9,1])) enddef
startdef d = random.choice([55,60,65,70,75,80]) enddef

startdef zerospeed = round((insvar*d)**.5,2) enddef
startdef slimit = round((zerospeed)/5)*5+random.choice([-10,-5,5,10]) enddef %some hit, some don't, going the speed limit
startdef s = max(int(zerospeed),int(slimit)) + np.random.randint(3,15) enddef

startdef skidd = round(s*s/insvar,3) enddef
startdef safed = round(slimit*slimit/insvar,3) enddef
startdef rsafed = int(safed) enddef
%%%%



startpre

endpre


startpost

endpost




startall
startquestion 
\item Solve each and include units in your answer. Use the formula: $s = \sqrt{\insvar \cdot d}$ where $s$ is the speed of the car in miles per hour prior to braking, and $d$ is the stopping distance or length of the skid mark, in feet. 

\vspace{3mm}

Deone is driving down Bumpkin Road going $\slimit$ miles per hour when a fawn suddenly appears $\d$ feet away in the middle of the road. \begin{enumerate}
\item {\bf (5 points)} If she slams on her brakes now, how far will her car skid? \vspace{4cm}
\item {\bf (2 points)} Will she avoid hitting the fawn if it freezes in place? Why or why not? Fully explain your reason. \vspace{3cm}
\end{enumerate}
endquestion 

startsolution
\item Formula $s = \sqrt{\insvar \cdot d}$. Deer is $\d$ feet away. \begin{enumerate}
\item If speed is $\slimit$ miles per hour, she will skid $\safed$ feet.
\item She will \ifthenelse{\rsafed > \d}{hit the deer}{not hit the deer}.
\end{enumerate}
endsolution
endall